---
title: "Machine Learning with modelr and caret"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document provides the necessary syntax...

Important Note: this tutorial is based on the introductions to modelr and caret, created by Hadley Wickham and Max Kuhn, respectively. (I have copied their basic layouts along with a number of specific examples.) Additional information can be found at: https://github.com/tidyverse/modelr. If you have a continued interest in using R for statistical/machine learning or would like additional information on the caret package I highly recomend the book "Applied Predictive Modeling"

Before we begin you should install and load the following packages:
```{r message=FALSE}
library(tidyverse)
library(modelr)
library(caret)
library(mlbench)
```

## Data 

```{r}
data("cars", package = "caret")
glimpse(cars)
cars <- as_tibble(cars)
```


## Data partitioning

The modelr package utilizes the `resample` class to create data partitions/samples. A `resample` object simply stores integer values that point back the the rows of the original data. Note: this can save a ton of space! We can create a resample object using `resample()`, view its pointer values by calling `as.integer()` or convert it to a data frame.

```{r}
rs <- resample(cars, 1:10)
as.integer(rs)
as.data.frame(rs)
```

We can easily create training, test or validation sets using `resample_partition()`

``` {r}
ex <- resample_partition(cars, c(test = 0.3, train = 0.7))
ex
```

We can manully fit a model by refering to the training data `ex$train`

``` {r}
mod <- lm(Price ~ Mileage + Doors + Cylinder, data = ex$train)
summary(mod)
```

We could then make predictions using the model

```{r}
predict(mod,ex$train) %>% 
  head()
```

```{r}
test <- ex$test %>% 
  as_tibble()
```

```{r}
test %>% 
  mutate(fit=predict(mod,test)) %>% 
  glimpse()
```

Manually computing performance

```{r}
test %>% 
  mutate(pred=predict(mod,test)) %>% 
  summarise(rmse=sqrt(mean((Price-fit)^2)))
```

Helper function in modelr

```{r}
test %>% 
  add_predictions(mod) %>% 
  glimpse()
```


``` {r}
rmse(mod,test)
rmse(mod,ex$test)
rsquare(mod,ex$test)
mae(mod,ex$test)
```


## Cross-Validation using map function

```{r}
folds <- crossv_kfold(cars, k = 5)
```

We now have to use the map function to apply a model to each training set

```{r}
folds <- folds %>% 
  mutate(model=map(train,~lm(Price ~ Mileage + Doors + Cylinder, data=.)))
```

```{r}
folds %>% 
  mutate(insamp_rmse=map2_dbl(model,train,~rmse(.x,.y))) %>% 
  mutate(outsample_rmse=map2_dbl(model,test,~rmse(.x,.y)))
```

``` {r}

```


## Parsimony with caret

The statistical/machine learning environment in r is strung out across a variety of different packages, with many utilizing a unique syntax and set of data structures. The caret pacakge attemps to synchronize these packages using a consistent syntax.

The 

``` {r message=FALSE}
data(Sonar)
ctrl <- trainControl(method="cv", 
                     summaryFunction=twoClassSummary, 
                     classProbs=T,
                     savePredictions = T)

set.seed(1234)
rf_fit <- train(Class ~ ., 
                data=Sonar, 
                method="rf", 
                preProc=c("center", "scale"), 
                trControl=ctrl)
```

## Go Parallel

```{r}

```


